### 2018-06-25 latex/results.tex
updated results

As our simulations suggest that NNMT might have excerted an evolitionary pressure on the development of NamPRT. We performed multiple sequence analysis to see if we can find sequence variations in the protein sequence of NamPRT that indicate evolutionary changes of NamPRT upon the appearance of NNMT. The multiple sequence alignment of selected sequences is shown in Figure 5A and Supplementary Figure ?\todo[author=Ines]{Decide figure number} We recognized that Deuterostomia that have only NamPRT and NNMT (indicated by the blue circle) have an insert of ten amino acids corresponding to positions 43 to 52 of the human enzyme, that overlaps with a predicted weak nuclear localisation signal, that is lost when removing this sequence stretch.
Furthermore, the inserted sequence corresponce to an structurallly unresolved loop structure in all avaibable crystal structure of human NamPR (e.g. \cite{Wang2006} structure visualisation Figure 5B). The loop is connected to one of the beta-sheets involved in substrate binding.

From these observations we derived two hypotheses regarding the potential function of the loop for NamPRT function. Firstly, we analysed whether the deletion of the amino acids 43 to 52 has an effect on the localisation of the human enzyme. But both wildtype and mutant protein show a mixed localisation in the cytosol and nucleus (Figure 5C). We therefore conclude that the partial nuclear localisation of NamPRT is not affected by the sequence deletion.
The second hypothesis was that the sequence insertion influences NamPRT affinity and activity, which is also what our mathematical simulations would predict. We thus recombinantly expressed  both wildttype and mutant  NamPRT  in E. coli.  Upon purifictaion we meassured the activity using NMR detecttion of NMN produced in the presence and absence of ATP. It appears that the enzymatic activity of the mutant enzyme is much lower than that of the wildtype enzyme both with and without ATP. We furthermore see that the activity of the mutant enzyme can be increase with higher concentration of Nam and PRPP whereas that is not the case for the wildtype enzyme. Pointing to a decrease substrate affinity of the mutant enzyme. Supporting the second hypothesis and thus supporting the predictions derived from our mathematical modelling approach.

@@

But why is NADA lost in vertebrates? Our simulations show that in direct competition NADA is outcompeted by the combined presence of NamPRT and NNMT but not by NamPRT alone. This effect deisappears of NamPRT affinity is lowered. ....

>>>>

As our simulations suggest that NNMT might have excerted an evolitionary pressure on the development of NamPRT. We performed multiple sequence analysis to see if we can find sequence variations in the protein sequence of NamPRT that indicate evolutionary changes of NamPRT upon the appearance of NNMT. The multiple sequence alignment of selected sequences is shown in Figure 5A. A more comprehensive multiple sequence alignment containing a larger number of species can be found in Supplementary Figure 1.  We recognized that most Deuterostomia that have only NamPRT and NNMT (indicated by the blue circle) have an insert of ten amino acids corresponding to positions 43 to 52 of the human enzyme. This insert overlaps with a predicted weak nuclear localisation signal, that is lost when removing this sequence stretch. The inserted sequence corresponce to a structurallly unresolved loop structure in all avaibable crystal structure of human NamPR (e.g. \cite{Wang2006} structure visualisation Figure 5B). The loop is connected to one of the beta-sheets involved in substrate binding.

From these observations we derived two hypotheses regarding the potential function of the loop for NamPRT function. Firstly, we analysed whether the deletion of the amino acids 42 to 51 has an effect on the localisation of the human enzyme. But both wildtype and mutant protein show a mixed localisation in the cytosol and nucleus (Figure 6C). We therefore conclude that the partial nuclear localisation of NamPRT is not affected by the sequence deletion.
The second hypothesis was that the sequence insertion influences NamPRT affinity and activity, which is also what our mathematical simulations would predict. We thus recombinantly expressed  both wildttype and mutant  NamPRT  in E. coli.  Upon purifictaion we meassured the activity using NMR detecttion of NMN produced in the presence and absence of ATP. It appears that the enzymatic activity of the mutant enzyme is much lower than that of the wildtype enzyme (Figure 6D). We furthermore see that in contrast to NamPRT wildtype, the mutant does not show an increased activity in the presence of ATP (Figure 6E).

@@

But why is NADA lost in vertebrates while NADA and NamPRT are co-existing existing in bacteria? To analyse this question we built a two compartment model that has a shared Nam source (details see Materials an Methods and Supplementary Material). One compartment contains NADA while the other contains either NamPRT alone or together with NNMT.  Without NNMT the compartment containing NADA has slightly lower NAD consumption rates (Figure 7A), but is able to maintain much higher steady state NAD concentrations at low cell division rates  (Figure 7B). At high cell division rates, however, steady state concentration in both compartments are similar. The latter might explain why in bacteria that have relatively high growth rates both systems coexists. 
If in contrast, NNMT is present in the NamPRT compartment, this compartment outcompetes the NADA containing compartment, both with respect to NAD-consumption rates and steady state NAD concentrations (Figure 7 C and D). This effect is dependent on a high affinity of NamPRT, as at low affinity the NADA containing system again maintains higher NAD-concentrations, the NAD consumption flux still being higher  in the NamPRT and NNMT containing compartment. Taken together this might explain why upon the development of a sufficiently high affinity of NamPRT for its substrate, which seems to have been induced by the appearance of NNMT, NADA is lost.
